---
title: "Real_Mini_Project_3"
author: "Paige E. Patrick"
date: "11/17/2017"
output: html_document
---

```{r}
library(mdsr)
library(RMySQL)
db <- dbConnect_scidb(dbname = "imdb")
```

```{r}
library(ggplot2)
library(tidyverse)
library(dbplyr)
```

```{r}
this_big_thing <- db %>%
dbGetQuery("SELECT mi.info, 
t.title, t.production_year, t.id
FROM title t
JOIN movie_info mi ON t.id = mi.movie_id
WHERE mi.info_type_id = 3
AND t.kind_id = 1;")
```

```{r}
movie_genre <- this_big_thing %>%
  group_by(info) %>%
  summarise(n_obvs = n())
```

```{r}
title_genre <- this_big_thing %>%
  group_by(production_year, id) %>%
  summarise(genres_observed = n()) %>%
  filter(production_year < 2018)
```

```{r}
genre_count_pre_1947 <- this_big_thing %>%
  group_by(production_year, id) %>%
  summarise(genres_observed = n()) %>%
  filter(production_year < 1947)
```

```{r}
genre_count_post_1947 <- this_big_thing %>%
  group_by(production_year, id) %>%
  summarise(genres_observed = n()) %>%
  filter(production_year > 1946, production_year < 2018)
```
<<<<<<< HEAD
```{r}

ggplot(movie_genre, aes(x=n_obvs, y=info)) + geom_point(shape=1)
```


=======

```{r}
avg_genre_count <- title_genre %>%
  group_by(production_year) %>%
  summarise(average_genre_count = mean(genres_observed))
```

```{r}
max_genre_count <- title_genre %>%
  group_by(production_year) %>%
  summarise(max_genre_count = max(genres_observed))
```

```{r}
min_genre_count <- title_genre %>%
  group_by(production_year) %>%
  summarise(min_genre_count = min(genres_observed))
```

```{r}
ggplot(avg_genre_count, aes(x =production_year, y = average_genre_count)) +
geom_line()
```

```{r}
ggplot(max_genre_count, aes(x = production_year, y = max_genre_count)) +
geom_line()
```

```{r}
<<<<<<< HEAD
avg_max_genre_count <- title_genre %>%
  group_by(production_year) %>%
  summarise(max_genre_count = max(genres_observed), average_genre_count = mean(genres_observed))
```
```{r}
ggplot(avg_max_genre_count, aes()) + 
  geom_line(aes(x = production_year, y = average_genre_count)) + 
  geom_line(aes(x = production_year, y = max_genre_count))
```
=======
genre_breakdown <- this_big_thing %>%
  group_by(production_year, info) %>%
  summarise(genres_observed = n()) %>%
  filter(production_year < 2018)
```

```{r}
ggplot(genre_breakdown, aes(x = production_year, y = genres_observed)) +
geom_line() +
  facet_wrap("info")
```

```{r}

```








>>>>>>> bbdaec12899b84c243adf753bdf0f2e5d0405ec9

```{r}

