---
title: "Genre Wars: IMDb Strikes Back"
author: "Paige E. Patrick & Vanessa Morales"
date: "12/3/2017"
output: html_document
---
According to Forbes, for every two dollars spent making a film, studios spend another dollar on marketing. Is this money all for nothing? IMDb genre categorization often conflicts with how a movie is marketed. We look at this disparity and its impact on how movies are perceived.

Production companies spend time and effort controlling how a movie is marketed, in order to attract the attention of certain audiences and increase sales for the movie. Because IMDb does not gain profit from movie sales, they do not need to be concerned with appealing to certain audiences. This allows IMDb to be more flexible with how they categorize movies on their website. We will be looking at a couple of the most popular movies over the past sixty years to see if IMDb has categorized them the same way they were marketed when they were first released. 

```{r, include = FALSE}
library(mdsr)
library(RMySQL)
db <- dbConnect_scidb(dbname = "imdb")
```

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(ggplot2)
library(tidyverse)
library(dbplyr)
library(knitr)
```


```{r}
sql_genre_data <- db %>%
dbGetQuery("SELECT t.title, mi.info, t.production_year, t.id
FROM title t
JOIN movie_info mi ON t.id = mi.movie_id
WHERE mi.info_type_id = 3
AND t.kind_id = 1;")
```

[We had done most of the heavy lifting for the code writing for the last project. This time was really just re-focusing the code we already had to help investigate our alternate hypothesis for the mini project three. We did this by filter the data to only include information on the four movies we were looking at. We picked these movies based off of the highest IMDb rated movies for each decade we are analyzing. This was as simple as looking up the information on the IMDb website.]

```{r}
movie_genre_info <- sql_genre_data %>%
  filter(id == 4361365 | id == 4118523 | id == 4260164| id == 3649367)
```

[The next step was being able to show the different genres for each of the four movies we are looking at.]

```{r}
dictator_info <- movie_genre_info %>%
  filter(id == 4361365) %>%
  select(info) 

psycho_info <- movie_genre_info %>%
  filter(id == 4118523) %>%
  select(info) 

starwars_info <- movie_genre_info %>%
  filter(id == 4260164) %>%
  select(info)

gladiator_info <- movie_genre_info %>%
  filter(id == 3649367) %>%
  select(info)
```  

[We didn't think that graphing the genres for each movie would be more beneficial to the audience than just showing it as is. A graph for information this straight forward seemed counter-productive. We used kable to create mini tables illustrating the genres in each of the movies. However, the caption was not showing up, so we had to create titles for each.] 



#####IMDb Genres recorded for The Great Dictator, 1940
```{r warning = FALSE, message = FALSE}
kable(dictator_info[1:3, ], caption = "The Great Dictator (1940)")
```

                                                
#####IMDb Genres recorded for Psycho, 1960
```{r warning = FALSE, message= FALSE}
kable(psycho_info[1:3, ], caption = "Psycho (1960)")
```


#####IMDb Genres recorded for Star Wars: The Empire Strikes Back, 1980
```{r warning = FALSE, message = FALSE}
kable(starwars_info[1:4, ], caption = "Star Wars: The Empire Strikes Back (1980)")
```


#####IMDb Genres recorded for Gladiator, 2000
```{r warning = FALSE, message = FALSE}
kable(gladiator_info[1:3, ], caption = "Gladiator (2000)")
```


[Editors Note: We will have four body paragraphs. Each will include the "IMDb Genres recorded" kable of the movie and a movie poster for each of these movies. We will discuss how the genre marketing for these movies does not match what IMDb has recorded for them. After we analyze each movie with its poster, we will have another kable showing the difference in how the movies were marketed. We did not add any of this stuff, since the rough draft was only focused on the baseline points (title, lede, and nutgraph). We will also hide the code we used for the kable code, but we kept it here to show we used code to create these tables. We will use code folding after we get input on the rough draft.]  





  
  
  