---
title: "Genre Multiplicity"
author: "Paige Patrick and Vanessa Morales"
date: "11/17/2017"
output: html_document
---

```{r, include = FALSE}
library(mdsr)
library(RMySQL)
db <- dbConnect_scidb(dbname = "imdb")
```

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(ggplot2)
library(tidyverse)
library(dbplyr)
```

The biggest challenge during this project was finding the data necessary to answer the prompt in SQL and getting it into R Studio. However, with the help of the indexes in the title, movie info, and kind id tables, we were able to quickly determine the information we needed to take from each of these tables. For instance, the movie info index helped us identify the info kind type that genre is registered under.

```{r}
sql_genre_data <- db %>%
dbGetQuery("SELECT mi.info, 
t.title, t.production_year, t.id
FROM title t
JOIN movie_info mi ON t.id = mi.movie_id
WHERE mi.info_type_id = 3
AND t.kind_id = 1;")
```

Next we needed to wrangle the data so that each production year had a list of all the movies produced in that year with the number of genres each movie has recorded as having. This is also when we noticed that the production year exceeds present day, so we needed to filter that to not have any missing data.

```{r}
genre_count <- sql_genre_data %>%
  group_by(production_year, id) %>%
  summarise(genres_observed = n()) %>%
  filter(production_year < 2018)
```

Then we took the table listed above and used that as a way to find the average number of genres movies seemed to be recorded of having within each production year. We also found the movie within each production year that had the maximum number of genres recorded.  

```{r}
genre_max_and_avg_count <- genre_count %>%
  group_by(production_year) %>%
  summarise(max_genre_count = max(genres_observed), average_genre_count = mean(genres_observed))
```

We thought comparing the maximum number of genres recorded for a movie and the average recorded for all movies would be interesting. The graph helps illustrate the gap between the maximum amount of genres recorded for a single movie and the average recorded for all movies in that year.

```{r, warning = FALSE}
ggplot(genre_max_and_avg_count, aes()) + 
  geom_line(aes(x = production_year, y = max_genre_count), colour = "green4", size = 1.5) +
  geom_line(aes(x = production_year, y = average_genre_count), colour = "magenta4", size = 1.5) +
  ggtitle("Genre Multiplicity in Movies,\nRecorded By IMDb") +
  scale_x_continuous("Production Year", breaks = c(1880, 1900, 1920, 1940, 1960, 1980, 2000, 2020)) +
  scale_y_continuous("Number of Genres Recorded", breaks = c(2, 4, 6, 8, 10, 12, 14)) +
  annotate("text", x = 2010, y = 3, label = "AVERAGE", colour = "magenta4", size = 6, face = "bold") +
  annotate("text", x = 2010, y = 6, label = "MAXIMUM", colour = "green4", size = 6, face = "bold") +
  theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold"), 
    axis.ticks.length = unit(0.25, "cm"),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13),
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 15),
    panel.background = element_rect(fill = "cadetblue2"),
    panel.grid.major = element_line(colour = "cadetblue3"),
    panel.grid.minor = element_line(colour = "cadetblue3"))
```

We also thought it would be interesting to look at if any genres on their own are maybe more popular now than in the first half of the century we are looking at. For some reason there was a genre labeled "Reality-TV" and "Reality-tv". Because "Reality-tv" had a very small number of points, we excluded it from the graph to keep from confusing the reader. Excluding the extremely small number of data points did not change the integrity of the data graphic. We checked.

```{r}
genre_breakdown <- sql_genre_data %>%
  group_by(production_year, info) %>%
  summarise(genres_observed = n()) %>%
  filter(production_year < 2018, info == "Action" | info == "Adult"| info == "Adventure" | info == "Animation" | info == "Biography" | info == "Comedy" | info == "Crime" | info == "Documentary" | info == "Drama" | info == "Family" | info == "Fantasy" | info == "Film-Noir" | info == "Game-Show" | info == "History" | info == "Horror" | info == "Music" | info == "Musical" | info == "Mystery" | info == "News" | info == "Reality-TV" | info == "Romance" | info == "Sci-Fi" | info == "Short" | info == "Sport" | info == "Talk-Show" | info == "Thriller" | info == "War" | info == "Western")
```

We wanted to give a very basic, and easily comprehend able graphic showing the overall trends throughout the years for each genre. We did not think labeling the y-axis was necessary to understand the trends. This was a style preference.

```{r}
ggplot(genre_breakdown, aes(x = production_year, y = genres_observed)) +
geom_line(colour = "red4", size = 1) +
  ggtitle("Popularity of Genres,\nRecorded By IMDb") +
  scale_x_continuous("Year", breaks = c(1900, 1950, 2000)) +
  scale_y_continuous("Basic Genre Trend", labels = c("", "", "", "", "")) + 
  theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold"), 
    axis.ticks.length = unit(0.25, "cm"),
    axis.ticks.y = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    panel.background = element_rect(fill = "cadetblue2"),
        panel.grid.major = element_line(colour = "cadetblue2"),
        panel.grid.minor = element_line(colour = "cadetblue2")) +
  facet_wrap("info")
```



##Genre Multiplicity

We are interested in seeing if movies made today have more genres than those made in the past. Initially, we believed that current movies would have higher genre multiplicity, and the older the movie, the more likely it would fit squarely into one genre.  

In researching our above inquiry, we hypothesized that there would be more genres used today than in the past. However, our data wrangling showed an overall average number of genres recorded to be around two to three throughout the century. What we did find however, was the maximum number of genres recorded for a single movie each year has increased. We believe that experimental marketing has caused some films to be tagged with an unusually high number of genres, getting as high as 15, as shown on our graph. This is shown as the max and not the average. It is possible that films advertised with two genres were the most successful, and studios used that as a standard. It is likely that films with higher genre multiplicity would be challeging to market and confusing for audiences. More genres does not mean more money or more popularity. This would explain why the average number of genres has stayed consistent throughout the century.

We graphed the data to show the popularity of each movie genre, recorded by IMDb starting from 1874 to 2017. Most genre trends stay relatively consistent, such as Family, Action, and Romance. A notable exception is the Film-Noir genre. It relays a distinct and brief period of popularity, from 1944 to 1954. The beginning of the genre can be attributed to the start of a new trend and style that hadn't previously existed. In order for a movie to be considered a part of the Film-Noir genre, it must be in black and white. It is likely the moody black and white quality of noir films ceased to be popular after color motion picture films became popular in the mid 1950s, as well as the increasing popularity of television. 

The Talk-Show, Reality-TV, and Game-Show graphs all show a direct correlation with the increase in home televisions in households around the mid-1950s. These types of TV-specific genres would not exist prior to the invention of the television.  

Lastly, we notice a large uptick in the popularity of the Short movie genre. We believe that this might relate to people's decreasing attention spans, the popularity of YouTube, the fixation on Facebook, and other social media consumption. The Short film genre peaked in popularity around the time YouTube, Vimeo, Facebook, and other social media platforms became widely used. This trend might have an influence on how movies are made. Another explination for the increase in Short films in recent years is that the equipment and means for production in order to create a film are more obtainable than in the past. Cameras, editing software, and social media websites offer an easy way to distribute homemade or smaller production films.

After anaylizing the data and looking at the graphs, we began to think of how information is recorded into IMDb and how that impacts the results of the data for genre multiplicity. Someone from IMDb in recent years has gone back and recorded the genre information for films that are over 50, 60, 70 years old. This means that how the genres are recorded might not be an accurate representation of how they would have been categorized by someone within the year the film was produced. This leads us to our first alternate hypothesis: the data recording the genres for movies is not reflected in IMDb accurately. 
                                                        
We had a second alternate hypothesis that would explain why the data collected for IMDb did not reflect high genre multiplicity trend we thought we would find. We believe that the functionality of IMDb as a website limits its ability to register movies outside of the more commonly used genres. When we look at Netflix, we see hyper-personalized genres. The purpose of Netflix is to build personalized accounts for indiviual users. Therefore, these super complex and specific genres can exist because they are specific to very few Netflix users. However, IMDb needs to be able to be understood universally. Keeping the genres simple and easy to understand makes their website easier to navigate. According to the article in the Atlantic, "How Netflix Reversed Engineered Hollywood", Netflix uses 76,847 micro-genres or unique ways to describe types of movies. This allows for data-driven programming and personalized genres. By mixing data and intuition the genre-labeling of IMDb would be more specific, inclusive of multiple genres, and perhaps more accurately represent a movie. 





Git hub Repository for this project:
https://github.com/vmorales95/MiniProject3

WorksCited:
theAtlantic, "How Netflix Reverse Engineered Hollywood", Alexis C. Madrigal, 2014. 




